---
title: "Peer Assessment 1 for Reproducible Research"
output: html_document
---

## Loading and preprocessing the data

Load the data.

```{r echo=TRUE}
data <- read.csv(file="G:/Data Science Specialization/05 Research/W2Project/activity/activity.csv")

```

## What is the mean total number of steps taken per day

Calculate the total number of steps taken per day.

```{r, echo=TRUE}
library(plyr)
sumDailySteps <- ddply(data,.(date),summarize,steps=sum(steps,na.rm=TRUE))
sumDailySteps
```

Make a histogram of the total number of steps taken each day.

```{r, echo=TRUE}

library(ggplot2)
qplot(sumDailySteps$steps,
      geom="histogram",
      binwidth=1000,
      main="Histogram of steps taken daily",
      xlab="steps",
      xlim=c(0,25000)
      )
```

Calculate and report the mean and median of the total number of steps taken per day.

```{r, echo=TRUE}
mean(sumDailySteps$steps, na.rm=TRUE)
median(sumDailySteps$steps, na.rm=TRUE)
```


## What is the average daily activity patern

Make a time series plot of the interval (x-axis) and the average number of steps taken.

```{r, echo=TRUE}
meanDailySteps <- ddply(data,.(interval),summarize,steps=mean(steps,na.rm=TRUE))
ggplot(data=meanDailySteps, 
       aes(x=interval, y=steps)
       ) + geom_line() + 
  labs(title="Average number of steps taken by interval ")
```

Which of the intervals contains the maximum number of steps?

```{r, echo=TRUE}

meanDailySteps[which.max(meanDailySteps$steps),]
```

## Imputing missing values

Calculate and report the total number of missing values in the dataset.

```{r, echo=TRUE}
table(is.na(data$steps))
```

Replace missing values in the dataset with the mean steps for that interval.

```{r, echo=TRUE}
avgSteps <- meanDailySteps$steps[match(data$interval,meanDailySteps$interval)]
```

Create a new dataset that is equal to the original but with the missing data filled in.

```{r, echo=TRUE}
df <- data
df$steps <- ifelse(is.na(df$steps),avgSteps,df$steps)
```

Make a histogram of the total number of steps taken each day.

```{r, echo=TRUE}
sumDailySteps2 <- ddply(df,.(date),summarize,steps=sum(steps,na.rm=TRUE))
qplot(sumDailySteps2$steps,
      geom="histogram", 
      binwidth=1000,
      main="Histogram of steps taken daily",
      xlim=c(0,25000),
      xlab="steps")
```

Calculate and report the mean and median total number of steps taken per day

```{r, echo=TRUE}
mean(sumDailySteps2$steps)
median(sumDailySteps2$steps)
```

The mean and median values now differ from the estimates from the first part of the assignment.

```{r, echo=TRUE}
mean(sumDailySteps$steps)!=mean(sumDailySteps2$steps)
median(sumDailySteps$steps)!=median(sumDailySteps2$steps)
```

The mean and median estimates are now higher after imputing missing data.

```{r, echo=TRUE}
mean(sumDailySteps2$steps)>mean(sumDailySteps$steps)
median(sumDailySteps2$steps)>median(sumDailySteps$steps)
```

## Are there differences in the activity patterns between weekdays annd  weekends

Create a new factor variable in the dataset indicating whether a given date is a weekend or weekday.
```{r echo=TRUE}

df$weekend <- weekdays(as.Date(df$date))%in%c("Saturday","Sunday")
df$weekend <- ifelse(df$weekend,"weekend","weekday")
```

Make a panel plot containing a time series plot of the interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days.

```{r echo=TRUE}
meanDailySteps2 <- ddply(df,.(interval,weekend),summarize,steps=mean(steps,na.rm=TRUE))

ggplot(meanDailySteps2, aes(interval,steps)) + 
  geom_line() + 
  facet_grid(weekend~.)

```
